---
title: "ReproducibleResearch-Project1"
author: "Hari Ravichandran"
date: "May 4, 2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r libraries}
library(lubridate) #For date conversion
library(ggplot2) #For Plotting
library(gridExtra) #For qplot() -> grid.arrange()
```

# Load the Data
```{r load}
S <- read.csv("activity.csv") #Loads data into data frame
S$date <- ymd(S$date) #Convert 'date' column from factor to date
class(S$date) #Check class of S$date

S1 <- na.omit(S) #Remove all NA values
head(S1) #Print out S1
```

# Make Histogram of Daily Data
```{r histogram}
S_sum <- aggregate(steps~date, S1, sum) #Total Steps taken per day
head(S_sum)

#Histogram of total steps taken each day
qplot(steps, data = S_sum , geom = "histogram", binwidth = 500, fill = "red") + 
    labs(title = "Total Number of Steps Each Day", x = "Number of Steps", y = "Frequency")
```

## Compute Mean and Median
```{r meanmedian}
mean_daily <- mean(S_sum$steps) #Mean of total number of steps
print(mean_daily)

median_daily <- median(S_sum$steps) #Mdian of total number of steps
print(median_daily)

```

#Make Time-Series Plot Based on Intervals
```{r intervals}
S_intervals <- aggregate(steps~interval, S1, sum) #Sum up values at each interval
head(S_intervals)

plot(S_intervals$interval, S_intervals$steps, type = "l") #Make time-series plot

max_ind <- which.max(S_intervals$steps) #Find & print the max interval with the most steps
max_interval <- S$interval[max_ind]
print(max_interval)
```

# Total Number of NA Values in Data Set
```{r NA}
print(sum(is.na(S)))
```

## Fill in NA Values with Zeros for 5 Minute Interval
```{r FiveMin}
S_fill <- S
S_fill[is.na(S_fill)] <- 0 #We fill in the NA values with zero

head(S_fill)
```

# Make Histogram of Daily Data for Filled Data
```{r histogram-fill}
S_sum_fill <- aggregate(steps~date, S_fill, sum) #Use aggregate to sum up values at each time --> daily
head(S_sum_fill)

#Plot Histogram
qplot(steps, data = S_sum_fill , geom = "histogram", binwidth = 500, fill = "red")  + 
    labs(title = "Total Number of Steps Each Day", x = "Number of Steps", y = "Frequency")
```

## Compute Mean and Median for Filled Data
```{r meanmedian-fill}
mean_daily_fill <- mean(S_sum_fill$steps)  #Mean of total number of steps
print(mean_daily_fill)

median_daily_fill <- median(S_sum_fill$steps)  #Median of total number of steps
print(median_daily_fill)
```

# Differences between Weekdays and Weekends?
```{r weekdays}
W <- wday(S_fill$date) #Use lubridate wday()
Is_Weekday <- ifelse(W %in% c(2,3,4,5,6), "Weekday", "Weekend") #2 - 6 are Mon - Fri

S_fill$Is_Weekday <- factor(Is_Weekday) #Create the factor variable for weekdays/weekends

head(S_fill)
```
## Panel Plot with Time Series Plot for Weekdays/Weekends
```{r timeseries-weekdays}
split_wdays <- split(S_fill, S_fill$Is_Weekday)

wdays <- aggregate(steps~interval, split_wdays$Weekday, sum)
wknds <- aggregate(steps~interval, split_wdays$Weekend, sum)

q_wdays <- qplot(x = interval, y = steps, data = wdays, geom = "line") + ggtitle("Weekdays")
q_wknds <- qplot(x = interval, y = steps, data = wknds, geom = "line") + ggtitle("Weekends")

grid.arrange(q_wdays, q_wknds, ncol = 1)
```
